---
title: "Assignment 0"
subtitle: "Statistics 32950 (Spring 2023)"
author: "Ki Hyun"
date: "Due: 23:00 (CT) 2023-03-24"
output: pdf_document
---

## 1.

$$
r = \frac{\sum_{i = 1}^n \left(x_i - \bar{x}\right) 
\left(y_i - \bar{y}\right)}
{\sqrt{\sum_{i = 1}^n \left(x_i - \bar{x}\right)^2
\sum_{i = 1}^n \left(y_i - \bar{y}\right)^2}}
$$

Here,

$$
\bar{x} = \frac{1}{n} \sum_{i = 1}^n x_i
$$

Moreover,

$$
\bar{y} = \frac{1}{n} \sum_{i = 1}^n y_i
$$

\newpage

## 2.

```{r q2}
x = runif(30); y = x^3 + rnorm(30)/3
plot(y, x); abline(lm(y ~ x))
title("Cubic relationship in Linear")
```

\newpage

## 3.

$$
A = \left[\begin{matrix} a_1 & a_2 & a_3
\end{matrix} \right] =
\left[\begin{matrix} 3 & -4 & -1 \\
-6 & 0 & 5 \\
4 & 5 &7
\end{matrix} \right]
$$

$$
B = \left[\begin{matrix} b_1 & b_2 & b_3
\end{matrix} \right] =
\left[\begin{matrix} 1 & 0 & 0 \\
0 & 3 & 0 \\
0 & 0 & -2
\end{matrix} \right]
$$

$$
E = \left[\begin{matrix} e_1 & e_2 & e_3
\end{matrix} \right] =
\left[\begin{matrix} 1 & 0 & 0 \\
0 & 0 & 1 \\
0 & 1 & 0
\end{matrix} \right]
$$

### (a)

#### i)

When $A$ is multiplied by $B$ from the left (i.e., $B A$), the following row
operations on $A$ occurs:

- The first elements of each column remain the same.
- The second elements of each column gets multiplied by $3$.
- The third elements of each column gets multiplied by $-2$.

#### ii)

When $A$ is multiplied by $B$ from the right (i.e., $A B$), the following column
operations on $A$ occurs:

- The first column remains the same. ($a_1$)
- The second column gets multiplied by $3$. ($3a_2$)
- The third column gets multiplied by $-2$. ($-2a_3$)

#### iii)

When $A$ is multiplied by $E$ from the left (i.e., $E A$), the following row
operations on $A$ occurs:

- The first row remains the same.
- The second row becomes the third row.
- The third row becomes the second row.

#### iv)

When $A$ is multiplied by $E$ from the right (i.e., $A E$), the following column
operations on $A$ occurs:

- The first column remains the same.
- The second column becomes the third column. ($a_3$)
- The third column becomes the second column. ($a_2$)

### (b)

$$
v = \left[\begin{matrix} 7 \\ 3 \\ 24
\end{matrix} \right]
$$

#### i)

```{r 3_b_1}
A = matrix(c(3, -4, -1, -6, 0, 5, 4, 5, 7), nrow = 3, byrow = T)
v = matrix(c(7, 3, 24), nrow = 3, byrow = T)
solve(A) %*% v
```

$$
v = 2 a_1 - a_2 + 3 a_3
$$


#### ii)

$$
v = 7 b_1 + b_2 - 12 b_3
$$

#### iii)

$$
v = 7 e_1 + 24 e_2 + 3 e_3
$$

\newpage

## 4.

### (a)

$$
\sum_{i = 1}^n k^3 = \left(\sum_{i = 1}^n k \right)^2
$$

### Proof by induction)

#### i)

Let's first check that the statement is true for $n = 1$.

$$
1^3 = (1)^2
$$

#### ii)

Let's assume that the statement is true for $m$, 
(i.e., $\sum_{i = 1}^m k^3 = \left(\sum_{i = 1}^m k \right)^2$).

Then for $m + 1$:

$$
\begin{aligned}
\sum_{i = 1}^{m + 1} k^3 &=
\sum_{i = 1}^{m} k^3 + (m + 1)^3 \\
&= \left(\sum_{i = 1}^m k \right)^2 + m^3 + 3m^2 + 3m + 1 \\
&= \left( \frac{m(m + 1)}{2} \right)^2 + m^3 + 3m^2 + 3m + 1 \\
&= \frac{m^4}{4} + \frac{m^3}{2} + \frac{m^2}{4} + m^3 + 3m^2 + 3m + 1 \\
&= \frac{1}{4} \left( m^4 + 6m^3 + 13m^2 + 12m + 4 \right) \\
&= \frac{1}{4} \left( m^2 + 3m + 2 \right)^2 \\
&= \left( \frac{(m + 1)(m + 2)}{2} \right)^2 \\
&= \left(\sum_{i = 1}^{m + 1} k \right)^2
\end{aligned}
$$

$$
\therefore
\sum_{i = 1}^n k^3 = \left(\sum_{i = 1}^n k \right)^2
$$

$$
Q.E.D.
$$

### (b)

Let,

$$
A = \left[\begin{matrix} 1 & 0 & 0 & 0 \\
1 & 0 & 0 & 0 \\
1 & 0 & 0 & 0 \\
1 & 0 & 0 & 0
\end{matrix} \right]
$$

and

$$
B = \left[\begin{matrix} 1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1
\end{matrix} \right]
$$

Then,

$$
AB = \left[\begin{matrix} 1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1 \\
1 & 1 & 1 & 1
\end{matrix} \right]
$$

However,

$$
BA = \left[\begin{matrix} 4 & 0 & 0 & 0 \\
4 & 0 & 0 & 0 \\
4 & 0 & 0 & 0 \\
4 & 0 & 0 & 0
\end{matrix} \right]
$$

$$
\therefore
AB \neq BA
$$

$$
Q.E.D.
$$

\newpage

## 5.

$$
U \sim Uniform[-1, 1]
$$

$$
X = U^2
$$

### (a)

For $0 < x \leq 1$

$$
\mathbf{P}\{X \leq x\} =
\mathbf{P}\{-\sqrt{x} \leq U \leq \sqrt{x}\} =
\frac{2\sqrt{x}}{2} = \sqrt{x}
$$

Here,

$$
\frac{\delta}{\delta x} \mathbf{P}\{X \leq x\} =
\frac{\delta}{\delta x} \sqrt{x} =
\frac{1}{2\sqrt{x}}
$$

Therefore, the p.d.f. of $X$ ($f(x)$) can be described as:

$$
f(x) = \begin{cases}
\frac{1}{2\sqrt{x}} & 0 < x \leq 1 \\
0 & \text{otherwise}
\end{cases}
$$

### (b)

$$
\begin{aligned}
\mathbf{E}\left(X\right) &=
\int_0^1 x f(x) dx \\
&= \int_0^1 \frac{x}{2\sqrt{x}} dx \\
&= \int_0^1 \frac{\sqrt{x}}{2} dx \\
&= \left[ \frac{1}{3} x^{\frac{3}{2}}\right]_0^1 \\
&= \frac{1}{3}
\end{aligned}
$$